\bf An observational study of papers published in \emph{PLOS ONE} and studies posted to a trial registry.

### 2021-04-22 manuscript/asa_template/asa_template.Rmd
update asa version of manuscript

authors: 
- name: Nicole White
  thanks: The authors gratefully acknowledge ...
  affiliation: School of Public Health and Social Work, QUT

@@

We used two openly available data sources to find statistical methods sections, covering peer-reviewed journal articles and registered clinical trial protocols.

@@

All _PLOS ONE_ articles are freely accessible via the PLOS Application Programming Interface (API). This enabled us to conduct semi-automated searches of full-text articles and analyse data on individual records, including text content and general attributes such as publication date and field(s) of research. To find papers with a statistical methods section we used targeted API searches followed by article filtering based on section headings. The data were downloaded on 3 July 2020.

@@

_Step 2_: Partial matching on section headings. Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s). Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings agai

@@

nst frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’. To determine the reliability of our chosen filters, we manually reviewed full text data extracted for a random sample of XXX articles that were not matched (File S1).[TODO…finish this thought…]

@@

@@

Results were examined for evidence of boilerplate text in two ways. We first reviewed the top ten results that represented the strongest matches to each cluster. To compare results within a cluster, we calculated cosine similarity score for all pairs of records; higher scores denoted a higher degree of similarity in text between a pair of records.

@@

```{r, echo=FALSE, fig.align='center', out.width="95%", fig.cap="PLOS ONE result summary."}
knitr::include_graphics("figures/plos.summary.jpg")
```

@@

```{r,fig.height=12,fig.width=6,fig.asp=0.55,fig.cap='\\label{fig:plos-n}Search results by _PLOS ONE_ volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}

@@

API searches returned `r format(n.api,big.mark=',')` unique records, of which `r format(n.include,big.mark=',')` (`r perc.include`%) included a statistical methods section based on our search criteria. In the final sample, 95,518 (85%) DOIs returned an exact match against common section headings, including 64,133 for 'statistical analysis', 13,380 for 'statistical analyses' and 13,627 for 'data analysis'. Among DOIs that did not meet the partial matching criteria, initial search terms appeared in [TODO].

@@

```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="PLOS ONE topic clouds."}
knitr::include_graphics("figures/plos.wordclouds.jpg")
```

We applied the clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50. Increasing the number of clusters decreased cluster quality based on global goodness-of-fit measures (Supplementary Figure 1), with average silhouette score and within-cluster dispersion leveling off around 20 clusters. This indicated that the data comprised one large, heterogeneous cluster and multiple smaller clusters.

```{r}
load('../../results/plos.results.10topics.rda')

ftab.plos = matches %>%
  group_by(topic_id) %>% summarise(DOIs = n(),`Median word count` = median(words),Q1=quantile(words,.25),Q3=quantile(words,.75),.groups='drop')

flextable(ftab.plos)

The topic clouds based on ten clusters are in Figure~\ref{fig:plot-10-topics}. Frequently occurring words reflected the use of statistical software (Topics~3 and 5), descriptive statistics (Topic~6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics~1 and 9). There are also statistical methods sections associated with regression (Topic~2) and meta-analysis (Topic~7).

@@

We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~X. 
Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.

>>>>

authors: 
- name: Nicole White
  thanks: The authors gratefully acknowledge computational resources and services used in this work provided by the eResearch Office, Queensland University of Technology, Brisbane, Australia.
  affiliation: School of Public Health and Social Work, QUT

@@

We used two openly available data sources to find statistical methods sections: research articles published in _PLOS ONE_ and study protocols registered on the Australian and New Zealand Clinical Trials Registry (ANZCTR). Data sources were chosen as examples of common research outputs that include descriptions of statistical methods that were used, or are planned, for analysing study outcomes.

@@

All _PLOS ONE_ articles are freely accessible via the PLOS Application Programming Interface (API). This enabled us to conduct searches of full-text articles and analyse data on articles' text content and general attributes such as publication date and field(s) of research. To find papers with a statistical methods section we used targeted API searches followed by article filtering based on section headings. The data were downloaded on 3 July 2020.

@@

_Step 2_: Partial matching on section headings. Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s). Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’. For records that did not pass this second stage filter, we selected a random sample of XXX records and reviewed where initial search terms appeared in the full-text, to estimate the likely proportion of statistical methods sections that were missed.

<!-- For records that did not pass this second stage filter, we reviewed where keywords appeared in the full-text to ascertain the extent of missing statistical analysis sections. ...[TODO: finish this whilst draft is with coauthors]-->

@@

@@

We applied the clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.

@@

For each dataset, records assigned to invididual clusters were examined for evidence of boilerplate text in two ways. We first reviewed the top ten results that represented the strongest matches to each cluster. Records assigned to the same cluster were also compared by calculating pairwise cosine similarities; higher scores denoted a higher degree of similarity in text between a pair of records.

@@

```{r, echo=FALSE, fig.align='center', out.width="70%",fig.cap="\\label{fig:plos-n}PLOS summary"}

knitr::include_graphics("figures/plos.summary.jpg")
```

@@

```{r}

@@

API searches returned `r format(n.api,big.mark=',')` unique records, of which `r format(n.include,big.mark=',')` (`r perc.include`%) included a statistical methods section based on our search strategy. In the final sample, 95,518 (85%) DOIs returned an exact match against common section headings, including 64,133 for 'statistical analysis', 13,380 for 'statistical analyses' and 13,627 for 'data analysis'. Among DOIs that did not meet the partial matching criteria, initial search terms appeared in [TODO].

@@

```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="\\label{fig:plos-wordclouds}PLOS ONE topic clouds."}
knitr::include_graphics("figures/plos.wordclouds.jpg")
```

Increasing the number of clusters decreased cluster quality based on global goodness-of-fit measures (Supplementary Figure 1), with average silhouette score and within-cluster dispersion leveling off around 20 clusters. This indicated that the data comprised one large, heterogeneous cluster and multiple smaller clusters.

The topic clouds based on ten clusters are in Figure~\ref{fig:plos-wordclouds}. Frequently occurring words reflected the use of statistical software (Topics 3 and 5), descriptive statistics (Topic 6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics~1 and 9). Statistical methods sections associated with regression (Topic 2) and meta-analysis (Topic 7) were also identified.

<!-- ```{r} -->
<!-- load('../../results/plos.results.10topics.rda') -->

<!-- ftab.plos = matches %>% -->
<!--   group_by(topic_id) %>% summarise(DOIs = n(),`Median word count` = median(words),Q1=quantile(words,.25),Q3=quantile(words,.75),.groups='drop')  -->

<!-- flextable(ftab.plos) -->

<!-- ``` -->

@@

We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~\ref{fig:anzctr-missing}. Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.

```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="ANZCTR search results."}

knitr::include_graphics("figures/excluded_anzctr_missing.jpg")

### 2021-04-22 manuscript/asa_template/asa_template.tex
update asa version of manuscript

  \author{
        Nicole White \thanks{The authors gratefully acknowledge \ldots{}} \\
    School of Public Health and Social Work, QUT\\
     and \\     Thiru Balasubramaniam, Richi Nayak \\
    To add\\
     and \\     Adrian Barnett \\
    School of Public Health and Social Work, QUT\\
      }
  \maketitle
} \fi

@@

We used two openly available data sources to find statistical methods
sections, covering peer-reviewed journal articles and registered
clinical trial protocols.

@@

All \emph{PLOS ONE} articles are freely accessible via the PLOS
Application Programming Interface (API). This enabled us to conduct
semi-automated searches of full-text articles and analyse data on
individual records, including text content and general attributes such
as publication date and field(s) of research. To find papers with a
statistical methods section we used targeted API searches followed by
article filtering based on section headings. The data were downloaded on
3 July 2020.

@@

\emph{Step 2}: Partial matching on section headings. Full text XML data
for all search results were downloaded and combined into a single
dataset, organised by DOI and subsection heading(s). Since \emph{PLOS
ONE} does not prescribe standardised headings to preface statistical
methods sections, we performed partial matching on available headings
agai

@@

nst frequently used terms in initial search results: `Statistical
analysis', `Statistical analyses', `Statistical method', `Statistics',
`Data analysis' and `Data analyses'. To determine the reliability of our
chosen filters, we manually reviewed full text data extracted for a
random sample of XXX articles that were not matched (File
S1).{[}TODO\ldots finish this thought\ldots{]}

@@

{\centering \includegraphics[width=0.95\linewidth]{figures/plos.summary}

@@

\caption{PLOS ONE result summary.}\label{fig:unnamed-chunk-3}
\end{figure}

API searches returned 131,847 unique records, of which 111,731 (85\%)
included a statistical methods section based on our search criteria. In
the final sample, 95,518 (85\%) DOIs returned an exact match against
common section headings, including 64,133 for `statistical analysis',
13,380 for `statistical analyses' and 13,627 for `data analysis'. Among
DOIs that did not meet the partial matching criteria, initial search
terms appeared in {[}TODO{]}.

@@

\caption{PLOS ONE topic clouds.}\label{fig:unnamed-chunk-5}
\end{figure}

We applied the clustering algorithm to the cleaned dataset, varying the
number of clusters from 1 to 50. Increasing the number of clusters
decreased cluster quality based on global goodness-of-fit measures
(Supplementary Figure 1), with average silhouette score and
within-cluster dispersion leveling off around 20 clusters. This
indicated that the data comprised one large, heterogeneous cluster and
multiple smaller clusters.

\includegraphics[width=3.75in,height=2.75in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-6-1.png}

The topic clouds based on ten clusters are in
Figure\textasciitilde{}\ref{fig:plot-10-topics}. Frequently occurring
words reflected the use of statistical software (Topics\textasciitilde3
and 5), descriptive statistics (Topic\textasciitilde6), group based
hypothesis testing (Topics 1 and 4) and definitions of statistical
significance (Topics\textasciitilde1 and 9). There are also statistical
methods sections associated with regression (Topic\textasciitilde2) and
meta-analysis (Topic\textasciitilde7).

@@

\includegraphics[width=6.13in,height=3.49in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-7-1.png}

@@

\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-8}
\end{figure}

We downloaded 28,008 studies. The numbers of excluded studies are shown
in Figure\textasciitilde X. Of the 12,700 included studies, 9,523 (75\%)
had a statistical methods section.

>>>>

  \author{
        Nicole White \thanks{The authors gratefully acknowledge computational resources and services
used in this work provided by the eResearch Office, Queensland
University of Technology, Brisbane, Australia.} \\
    School of Public Health and Social Work, QUT\\
     and \\     Thiru Balasubramaniam, Richi Nayak \\
    To add\\
     and \\     Adrian Barnett \\
    School of Public Health and Social Work, QUT\\
      }
  \maketitle
} \fi

@@

We used two openly available data sources to find statistical methods
sections: research articles published in \emph{PLOS ONE} and study
protocols registered on the Australian and New Zealand Clinical Trials
Registry (ANZCTR). Data sources were chosen as examples of common
research outputs that include descriptions of statistical methods that
were used, or are planned, for analysing study outcomes.

@@

All \emph{PLOS ONE} articles are freely accessible via the PLOS
Application Programming Interface (API). This enabled us to conduct
searches of full-text articles and analyse data on articles' text
content and general attributes such as publication date and field(s) of
research. To find papers with a statistical methods section we used
targeted API searches followed by article filtering based on section
headings. The data were downloaded on 3 July 2020.

@@

\emph{Step 2}: Partial matching on section headings. Full text XML data
for all search results were downloaded and combined into a single
dataset, organised by DOI and subsection heading(s). Since \emph{PLOS
ONE} does not prescribe standardised headings to preface statistical
methods sections, we performed partial matching on available headings
against frequently used terms in initial search results: `Statistical
analysis', `Statistical analyses', `Statistical method', `Statistics',
`Data analysis' and `Data analyses'. For records that did not pass this
second stage filter, we selected a random sample of XXX records and
reviewed where initial search terms appeared in the full-text, to
estimate the likely proportion of statistical methods sections that were
missed.

@@

@@

{\centering \includegraphics[width=0.7\linewidth]{figures/plos.summary}

@@

\caption{\label{fig:plos-n}PLOS summary}\label{fig:unnamed-chunk-3}
\end{figure}

API searches returned 131,847 unique records, of which 111,731 (85\%)
included a statistical methods section based on our search strategy. In
the final sample, 95,518 (85\%) DOIs returned an exact match against
common section headings, including 64,133 for `statistical analysis',
13,380 for `statistical analyses' and 13,627 for `data analysis'. Among
DOIs that did not meet the partial matching criteria, initial search
terms appeared in {[}TODO{]}.

@@

\caption{\label{fig:plos-wordclouds}PLOS ONE topic clouds.}\label{fig:unnamed-chunk-5}
\end{figure}

Increasing the number of clusters decreased cluster quality based on
global goodness-of-fit measures (Supplementary Figure 1), with average
silhouette score and within-cluster dispersion leveling off around 20
clusters. This indicated that the data comprised one large,
heterogeneous cluster and multiple smaller clusters.

The topic clouds based on ten clusters are in
Figure\textasciitilde{}\ref{fig:plos-wordclouds}. Frequently occurring
words reflected the use of statistical software (Topics 3 and 5),
descriptive statistics (Topic 6), group based hypothesis testing (Topics
1 and 4) and definitions of statistical significance
(Topics\textasciitilde1 and 9). Statistical methods sections associated
with regression (Topic 2) and meta-analysis (Topic 7) were also
identified.

@@

\includegraphics[width=6.13in,height=3.49in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-6-1.png}

@@

\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-7}
\end{figure}

We downloaded 28,008 studies. The numbers of excluded studies are shown
in Figure\textasciitilde{}\ref{fig:anzctr-missing}. Of the 12,700
included studies, 9,523 (75\%) had a statistical methods section.

{\centering \includegraphics[width=0.7\linewidth]{figures/excluded_anzctr_missing}

\caption{ANZCTR search results.}\label{fig:unnamed-chunk-8}
\end{figure}
